---
title: "Module3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Module3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(gbk346)
```


```{r}
mycalc_function <- function(x, y, fun) {
  if(fun == "sum") {
    o <- (x + y)
  } else if(fun == "substract") {
    o <- (x - y)
  } else if(fun == "multiply") {
    o <- (x * y)
  } else if(fun == "divide") {
    o <- (x / y)
  } else {
    o <- paste("Not a recognized function")
  }
  return(o)
}
```

##Creating Vectors
  -The following are vectors which will later be used later to test the mycalc_function.
```{r, echo=TRUE}
#a is the length of several vectors this is used as a hold over
a = 20

#v1
set.seed(1)
v1 <- sample(x = 1:100, size = a, replace = TRUE)

set.seed(1)
v2 <- sample(x = 1:10, size = a, replace = TRUE)

set.seed(1)
v3 <- runif(20, min = 0, max = 1)

v4 <- sample(x = LETTERS[v2], size = 20, replace = TRUE) 

```

##Creating Matrices
  -The following are matrices which will later be used to test the mycalc_function.
```{r}
m1 <- cbind(v1, v2)
m2 <- cbind(v2, v3)
m3 <- cbind(v3, v4)
```

## Creating Dataframes
  -The following is a dataframe which will later be used to test the mycalc_function.
```{r}
DF <- data.frame(v1, v2, v3, v4)
colnames(DF) <- c("a", "b", "c", "grp")

```


## Calculations 
  - The following are various different trial runs of the  mycalc_function.

```{r, error= TRUE}
s1 <- mycalc_function(v1, v2, "sum")
d1 <- mycalc_function(v1, v2, "divide")
s2 <- mycalc_function(v2, v3, "sum")
d2 <- mycalc_function(v2, v3, "divide")
s3 <- mycalc_function(v1, v3, "sum")
d3 <- mycalc_function(v1, v3, "divide")
s4 <- mycalc_function(m1, m1, "sum")
d4 <- mycalc_function(m1, m1, "divide")
s5 <- mycalc_function(m1, m2, "sum")
d5 <- mycalc_function(m1, m2, "divide")
s6 <- mycalc_function(m2, m3, "sum")
d6 <- mycalc_function(m2, m3, "divide")
```

## More Calculations

```{r, error= TRUE}
b1 <- mycalc_function(DF[1,1], DF[1,2], "multiply")
b2 <- mycalc_function(DF[1:10, 1:3], DF[1:10, 1:3], "multiply")
b3 <- mycalc_function(DF[DF$grp == "E", c("a", "b")], DF[DF$grp == "E", c("c")], "divide")
b4 <- (DF[1,1] * DF[1,2])
b5 <- DF[1:10, 1:3] * DF[1:10, 1:3]
b6 <- DF[DF$grp == "E", c("a", "b")] / DF[DF$grp == "E", c("c")] 
```

```{r, error= TRUE}
l <- list(v1, m1, m2)
myfunc <- lapply(1:length(l), function(x) mycalc_function(l[[x]], v2, "multiply"))
myfunc_2 <- sapply(1:length(l), function(x) mycalc_function(l[[x]], v2, "multiply"))
```

